---
title: "StudentworkBookWeek10"
author: "Student"
date: "11/12/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)

library(tidyverse)
library(readxl)
library(DBI)
library(knitr)
library(kableExtra)


#enable other connections
# dbDisconnect(con)
#
# connect to CapeCodd
con <- DBI::dbConnect(RSQLite::SQLite(), "hr.db")
#

#
# use the database connection for all SQL chunks in this notebook and hide the code by default
knitr::opts_chunk$set(connection = "con", echo=FALSE)
```

## Week 10 Assignment DDO First Semester

In this assignment we will be looking at the content studied during the DDO course. We will focus especially on the meaning of certain queries. Some questions require only a SQL code, others will ask you to explain in plain English how the code works. The same explanation will go for the database relations. 

>Being able to explain a set of code in simple wording is a good skill for IT engineers, regardless of their field of work. In your professional career, you will have to explain to a layman how a certain code is working. In most cases, this will be your manager or CEO, who has no IT experience. Commands like SELECT, WHERE and GROUP BY will not be understood and need additional explanation. 

you are allowed to work in groups, making this workbook also ensures that you have all libraries for the individual test in week 12. (HINT make sure each member of the group can make the assignments.)
Good Luck...

### Reading a DataBase

Looking at the ERD, shown in figure 1, we can see 6 different relations, describe all six in plain english going from 1 to many.

<img src = "images/week10.png" />

Relation 1 *Customer - Rental*: Mandatory One customer to Mandatory Many Rentals

Relation 2 *Rental - Employee*: Optional Many Rentals to Mandatory One Employee

Relation 3 *Rental - Room*: Optional Many Rental to Mandatory Many Rooms

Relation 4 *Hotel - Room*: Mandatory One Hotel to Mandatory Many Rooms

Relation 5 *Hotel - Employee (Employee)*: Mandatory One Hotel to Mandatory Many Employees

Relation 6 *Hotel - Employee (Managing)*: Optional One Hotel to Mandatory One Employee

## SQL statements

Make an SQL query for the following questions. The desired result is printed below each question. Additionally in some question you need to provide an explanation of the code in plain English. Asume you have to discuss your result with a layman and he wants to understand what you have done.

<img src = "images/HR.png" />

Make an SQL query for the following questions. The desired result is printed below each question. 

A.	The HR department needs a report to display the employee number, last name, salary, and salary increased by 15.5% for each employee with an employee ID below 110. Label the column New Salary.

```{sql, connection=con, echo=TRUE, error = TRUE }

SELECT employee_id, last_name, salary, ((salary * 115.5)/100) as New_salary
FROM employees
WHERE employee_id < 110;

```

B.	Find the highest, lowest, sum, and average salary of all employees. Label the columns Maximum, Minimum, Sum, and Average, respectively. Round your results to the nearest whole number.

```{sql, connection=con, echo=TRUE, error = TRUE }

SELECT Max(salary) as Maximum, Min(salary) as Minimum, Sum(salary) as SUM, Round(Avg(salary), 0) as Average
FROM employees;

```

C.  Explain in plain english the following code. 

```{sql connection=con, echo=TRUE}
SELECT job_id, MAX(salary) AS MaxSalary, MIN(salary) AS MinSalary, 
SUM(salary) AS SumSalary, avg(salary) AS AvgSalary 
FROM employees 
GROUP BY job_id;

```

 Your CEO wants to report highest, lowest, sum, and average salary of all employees for each specific job.


D. Write a query to display the number of employees with the same job.

```{sql, connection=con, echo=TRUE , error = TRUE}
SELECT job_id, COUNT(job_id) as total_number
FROM employees
GROUP BY job_id;

```

E. Determine the number of managers without listing them. Label the column Number of Managers. Hint: Use the MANAGER_ID column to determine the number of managers.

```{sql, connection=con, echo=TRUE, error = TRUE }

SELECT COUNT(DISTINCT(manager_id)) AS Number_of_Managers 
FROM employees;

```

F. Write a query to display the names (first_name, last_name, Hire_date, Job_id) using alias name "First Name", "Last Name", "Hired", "Job" for all employees who's first name starts with an "A"

```{sql, connection=con, echo=TRUE, error = TRUE }

SELECT first_name as First_Name, last_name as Last_Name, hire_date as Hired, job_id as Job  
FROM employees
WHERE first_name LIKE "A%";

```

G. Explain the following code in plain English

```{sql connection=con, echo=TRUE}
SELECT manager_id, MIN(salary) AS MinSalary 
FROM employees 
GROUP BY manager_id 
HAVING MIN(salary)> 6000 
ORDER BY  MinSalary DESC

```
Your team leader wants information for all managers (use their id ) with minimun salary (display it as MinSalary) bigger than 6000 starting from the biggest to smallest ( expected a DESC list).

H. Explain the following code in plain English, how could this chunk be improved using INNER JOIN?

```{sql connection=con,  echo=TRUE}
SELECT e.last_name, e.job_id, e.department_id, d.depart_name 
FROM employees e, departments d, locations l 
WHERE e.department_id = d.department_id 
AND d.location_id = l.location_id 
AND LOWER(l.city) = 'toronto';

SELECT e.last_name, e.job_id, e.department_id, d.depart_name
FROM employees e
INNER JOIN departments d
ON e.department_id = d.department_id
INNER JOIN locations l
ON d.location_id = l.location_id, LOWER(l.city) = 'toronto'
```
The HR department needs to find the last names, job title, department id and name for all employees who have marketing experience in Toronto.


I. The HR department needs to find the names, salary and hire dates for all employees who have a higher salary than their managers, along with their managers’ names and hire dates and salary.

```{sql, connection=con, echo=TRUE, error = TRUE }
SELECT e.last_name as employee, e.hire_date as Hired_Em, e.salary, m.last_name as Manager, m.hire_date as Hired_Ma, m.salary as Salary_Ma
FROM employees m
INNER JOIN employees e ON m.employee_id = e.manager_id AND m.salary < e.salary;

```

J. Explain the following code in plain English

```{sql connection=con, echo=TRUE}
SELECT last_name, department_id, job_id 
FROM employees 
WHERE department_id IN (
  SELECT department_id 
  FROM departments 
  WHERE location_id = 1700);
```
Your team leader needs specific information for location with id 1700 for all employees. He needs to display their last name, department id, and job id.

K. Create a report for HR that displays the last name and salary of every employee who reports to King.
```{sql connection=con, echo=TRUE, output.var = "df_result"}
SELECT e.last_name, e.salary
FROM employees e
WHERE manager_id = 100 OR manager_id = 156;
```

##BONUS

As with every bonus this is a more complext challange allowing to show more skills. 
It is still part of the whole assignment and as such taken into account with the formative indication. 

L. Make the following graph. (Hint, make sure to check the data set carefully, there are some hidden tricks)

```{sql, connection=con, echo=TRUE, error = TRUE }

```

--ggplot()
```{r, echo=TRUE}

```